### 2020-06-18 submission/manuscript.tex
Pat's first take at revision

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing data
generation and data analysis, we have found scientists feel empowered
when they can both ask and answer their own biological questions
{[}1{]}. Yet, the standard undergraduate and graduate training in the
biological sciences is typically insufficient for developing these
advanced data analysis skills. In our experience performing microbiome
research, it is more common to find exceptional bench scientists who are
inexperienced at analyzing large data sets than to find the reverse. Of
course this raises a challenge: How do we train bench scientists to
effectively answer biological questions with these larger datasets?

The ever-growing ability to generate data and feeling of unease in
analyzing the data are analogous to the struggles scientists also face
with engaging the voluminous scientific literature. A common strategy
for keeping up with the literature is participating in journal clubs,
which involve group discussion of a pre-selected paper. These range from
informal discussions to PowerPoint presentations and course credit. In
addition to staying current on the literature, journal clubs help
strengthen skills in critical thinking, communication, and integrating
the literature {[}2{]}. Our research group has leveraged the similarity
between the overwhelming nature of both scientific literature and data
analysis to address the challenge of teaching reproducible data analysis
practices to bench scientists. Over the past four years we have
experimented with a Code Club model to improve data analysis skills in a
community environment.

Our Code Club sessions are an hour long and alternate with Journal Club
as the second part of weekly two hour lab meetings. Initially, the Code
Club was used to review code from trainee projects. Instead of a
presentation, the presenter would project their code onto a screen and
the participants would go through the code, stating the logic behind
each line. This approach emphasized the importance of code readability
and gave beginners the opportunity to see the real-life, messy code of
more experienced peers. Unfortunately, the format only allowed us to
review a fraction of a project's code, making it difficult to integrate
the programmer's logic across their full project. A major issue with
this model was that sometimes beginners could not contribute to
improving the code, and even when they could, more experienced group
members would eventually dominate the discussion. This led to a lack of
participation by beginners, who would sometimes mentally check out and
resulted in an adversarial environment between the presenter and more
experienced members. As a result, presenters were reluctant to offer to
present again.

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}3{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using either R syntax from the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. Our Code Clubs
typically have 7 to 10 participants, but the inherent
``think-pair-share'' strategy should allow it to be scaled to larger
groups {[}4{]}.

We continue to experiment with approaches for running Code Club. During
the period of the Covid-19 pandemic in 2020 when many research labs were
shuttered, PDS started hosting an open virtual weekly Code Club.
Participants could engage the material posted to
\url{https://www.riffomonas.org/code_club/} via a Zoom meeting or
asynchronously using the recorded sessions available on YouTube.
Regardless of our approach to leading a Code Club, we have learned that
it is critical for the presenter to clearly articulate their goals and
facilitate participant engagement. Although some Code Club sessions may
be more experimental than others, on the whole they are a critical tool
to train bench scientists in reproducible data analysis practices. We
have provided some examples of successful topics in Table 1. We have
summarized the results of our own experimentation as Code Club
presenters and participants into Ten Rules. The first 3 rules apply to
all participants, Rules 4 through 8 are targeted to presenters, and the
last two for participants.

@@

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it often led to frustration for both
the presenter and participants. If the presenter shared their own code,
did they want participants to focus on their coding style or did they
want help incorporating a new package into their workflow? Participants
will always notice or ask about code concepts that are not the focus of
the exercise; a presenter with a specific goal can bring tangential
conversations back to the planned task. Where possible, presenters
should create a simplified scenario (i.e.~minimal, reproducible example
{[}5{]}), which can be helpful in focusing the participants. The
presenter should verify ahead of time that the simplified example works
and behaves the way they expect. Beyond the content, clear goals for
participant activities will help both parties stay on task and avoid
frustration. For more advanced learners, the presenter can create
stretch goals or give an activity with multiple stopping points where
participants would feel successful (e.g.~commenting code, creating
function, implementing function, refactoring function). Accomplishing
specific goals is more likely to result in a positive outcome for
presenters and participants.

@@

Participants want to learn topics that will either be useful to them or
help their colleague (i.e.~the presenter). Presenters should do their
best to satisfy those motivations, whether it is through the relevancy
of the concept or the data. It does not make sense to present a Code
Club on downloading stock market data if it is not useful or interesting
to the group. Similarly, participants should not be tasked with
improving the presenter's code if the presenter has no intention of
incorporating the suggestions. A list of packages or tasks that group
members are interested in could help a presenter struggling to find a
topic choose one that will make a rewarding Code Club.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. In these circumstances it can be useful to cover commonly
used tools but that might result in disinterest of more experienced
participants. We have identified several strategies to overcome the
challenges presented by participants at varying skill levels. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on their goals. Partnerships between those with similar skill
levels requires the presenter to design appropriate activities for each
skill level. We have found that commenting code is a good skill for
beginners since it forces them to dissect and understand a code chunk
line-by-line. It also reinforces the value of commenting as they develop
their skills and independence. An advantage of forming partnership
between people with disparate skill levels is that it is more likely for
groups to provide the presenter with a diverse range of methods that
achieve the same result. This approach to pairing also helps to graft
new members that have emerging programming skills. Regardless of how
partners are selected, consider asking the pairs to identify a navigator
and a driver {[}6{]}. The driver types at the computer while the
navigator tells them what to type, thus ensuring participation of both
partners. Midway through the activity, the presenter can have the
partners switch roles. Intentionally forming pairs can also engineer
group interactions by avoiding potentially disruptive partnerships or
pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion
about a paper that the participant has not read. In that context, coming
to Code Club without having installed a dependency is similar to asking
a simple question about the Journal Club paper. Both instances show a
lack of preparation. Just as a presenter must follow Rule 6 to provide
materials ahead of time, participants must review the code in advance,
download the data sets, install the necessary packages, and perhaps read
up on the topic. If the Code Club is based on a paper or chapters in a
data science book (e.g., {[}7--9{]}), the participants should read them
before the session and consider how they might incorporate the concepts
into their own work.

@@

\leavevmode\hypertarget{ref-Lonsdale2016}{}%
2. Lonsdale A, Penington JS, Rice T, Walker M, Dashnow H. Ten simple
rules for a bioinformatics journal club. Lewitter F, editor. PLOS
Computational Biology. 2016;12: e1004526.
doi:\href{https://doi.org/10.1371/journal.pcbi.1004526}{10.1371/journal.pcbi.1004526}

\leavevmode\hypertarget{ref-Wilson2014}{}%
3. Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et al.
Best practices for scientific computing. Eisen JA, editor. PLoS Biology.
2014;12: e1001745.
doi:\href{https://doi.org/10.1371/journal.pbio.1001745}{10.1371/journal.pbio.1001745}

\leavevmode\hypertarget{ref-Kothiyal2014}{}%
4. Kothiyal A, Murthy S, Iyer S. Think-pair-share in a large CS1 class.
Proceedings of the 2014 conference on innovation \& technology in
computer science education - ITiCSE 14. ACM Press; 2014.
doi:\href{https://doi.org/10.1145/2591708.2591739}{10.1145/2591708.2591739}

\leavevmode\hypertarget{ref-StackOverflow_MREE}{}%
5. Stack Overflow. How to create a minimal, reproducible example. 2020.
Available:
\url{https://stackoverflow.com/help/minimal-reproducible-example}

\leavevmode\hypertarget{ref-Bockeler_2020}{}%
6. Böckeler B, Siessegger N. On pair programming. 2020. Available:
\url{https://martinfowler.com/articles/on-pair-programming.html}

\leavevmode\hypertarget{ref-Wickham_2016}{}%
7. Wickham H, Grolemund G. O'Reilly Media; 2016. Available:
\url{https://r4ds.had.co.nz}

\leavevmode\hypertarget{ref-Schloss_2018}{}%
8. Schloss PD. The riffomonas reproducible research tutorial series.
Journal of Open Source Education. 2018;1: 13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}

\leavevmode\hypertarget{ref-Noble2009}{}%
9. Noble WS. A quick guide to organizing computational biology projects.
Lewitter F, editor. PLoS Computational Biology. 2009;5: e1000424.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000424}{10.1371/journal.pcbi.1000424}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.25\columnwidth}\raggedright
\textbf{Title}\strut
\end{minipage} & \begin{minipage}[b]{0.69\columnwidth}\raggedright
\textbf{Description}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.25\columnwidth}\raggedright
base vs.~ggplot2\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given input data and a figure, recreate the figure using R's base
graphics or ggplot2 syntax\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Snakemake\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a bash script that contains an analysis pipeline, convert it to a
Snakemake workflow (can also be done with GNU Make)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
DRYing code\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given script with repeated code, create functions to remove
repetition\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
mothur and Vegan\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a pairwise community dissimilarity matrix, compare communities
using the \texttt{adonis} function in the Vegan R package\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
tidy data\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a wide-formatted data table, convert it to a long, tidy-formatted
data table using tools from R's tidyverse\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
GitFlow\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Participants file and claim an issue to add their name to a README file
in a GitHub-hosted repository and file a pull request to complete the
issue\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
R with Google docs\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Scrape a Google docs workbook and clean the data to identify previous
Code Club presenters\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Develop an R Package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Convert a lab member's collection of scripts into an R package over a
series of sessions\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

>>>>

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing data
generation and data analysis, we have found scientists feel empowered
when they can both ask and answer their own biological questions
{[}1{]}. In our experience performing microbiome research, it is more
common to find exceptional bench scientists who are inexperienced at
analyzing large data sets than to find the reverse. Of course this
raises a challenge: How do we train bench scientists to effectively
answer biological questions with these larger datasets?

The standard undergraduate and graduate training in the biological
sciences is typically insufficient for developing these skills {[}2{]}.
To fill this void, there has been growth in online data science training
resources through companies including Data Camp
(\url{https://www.datacamp.com}) and Codecademy
(\url{https://www.codecademy.com}), massive open online courses (MOOCs),
and in person workshops hosted by organizations including The
Carpentries (\url{https://carpentries.org}). Each of these formats are
highly popular among their audience (e.g.~{[}3{]}); however, recent
analyses of workshops suggests that they have little impact on the long
term retention of the material that is taught {[}4{]}. This result
should not be surprising. These learning formats ask participants to
engage in massed practice; an approach that is known to be ineffective
{[}5--8{]}. In contrast, participants need regular opportunities to
engage in repeated deliberate practice that encourage them to practice
their retrieval and application skills of the material {[}9,10{]}. There
is clearly a need to develop a framework that can build upon material
introduced in coursework and workshops and that can help researchers to
stay up to date on the latest best practices, algorithms, and tools.

To address this need, we noted the similarity between the feeling of
unease in analyzing the data and the struggles scientists face with
engaging the voluminous scientific literature. A common strategy for
keeping up with the literature is by participating in a journal club.
Regardless of the discipline, a Journal Club involves group discussion
of a pre-selected paper. These range from informal discussions to
PowerPoint presentations and course credit. In addition to building upon
material from traditional coursework and staying current on the
literature, journal clubs help strengthen skills in critical thinking,
communication, and integrating the literature {[}11{]}. Since most
journal clubs occur on a regular schedule, they are effective because
they incorporate repeated deliberate practices. With this model in mind,
over the past four years we have experimented with creating a Code Club
model with the goal of improving reproducible data analysis skills in a
community environment.

Our Code Club sessions are an hour long and alternate with our lab's
Journal Club as the second part of weekly two hour lab meetings.
Initially, the Code Club was used to review code from trainee projects.
Instead of a presentation, the presenter would project their code onto a
screen and the participants would go through the code, stating the logic
behind each line. This approach emphasized the importance of code
readability and gave beginners the opportunity to see the real-life,
messy code of more experienced peers. Unfortunately, the format only
allowed us to review a fraction of a project's code, making it difficult
to integrate the programmer's logic across their full project. A major
issue with this model was that sometimes beginners could not contribute
to improving the code, and even when they could, more experienced group
members would eventually dominate the discussion. This led to a lack of
participation by beginners, who would sometimes mentally withdraw,
resulting in an adversarial environment between the presenter and more
experienced members. As a result, presenters were reluctant to offer to
present again.

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}12{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using either R syntax from the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. We generally
find that preparing a Code Club session takes as much effort as
preparing to present at a journal club. Our Code Clubs typically have 7
to 10 participants, but the inherent ``think-pair-share'' approach
should allow it to be scaled to larger groups {[}13{]}.

We continue to experiment with approaches for running Code Club. For
example, during the period of the Covid-19 pandemic in 2020 when many
research labs were shuttered, PDS started posting Code Club materials
(\url{https://www.riffomonas.org/code_club/}) that individuals or
research groups can use to practice their coding skills. During this
time, we continued our group's Code Club sessions by making use of the
breakout room and screen sharing features in Zoom. Regardless of our
approach to leading a Code Club, we have learned that it is critical for
the presenter to clearly articulate their goals and facilitate
participant engagement. Although some Code Club sessions may be more
experimental than others, on the whole they are a critical tool to train
bench scientists in reproducible data analysis practices. We have
provided some examples of successful topics in Table 1. We have
summarized the results of our own experimentation as Code Club
presenters and participants into Ten Rules. The first 3 rules apply to
all participants, Rules 4 through 8 are targeted to presenters, and the
last two for non-presenters. As the rules describe, this model can
easily be adapted to groups of non-biologists and to groups with
stronger background in computer science.

@@

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it often led to frustration for both
the presenter and participants. If the presenter shared their own code,
did they want participants to focus on their coding style or did they
want help incorporating a new package into their workflow? Participants
will always notice or ask about code concepts that are not the focus of
the exercise; a presenter with a specific goal can bring tangential
conversations back to the planned task. Where possible, presenters
should create a simplified scenario (i.e.~minimal, reproducible example
{[}14{]}), which can be helpful in focusing the participants. The
presenter should verify ahead of time that the simplified example works
and behaves the way they expect. Beyond the content, clear goals for
participant activities will help both parties stay on task and avoid
frustration. For more advanced learners, the presenter can create
stretch goals or give an activity with multiple stopping points where
participants would feel successful (e.g.~commenting code, creating
function, implementing function, refactoring function). Accomplishing
specific goals is more likely to result in a positive outcome for
presenters and participants.

@@

Participants want to learn topics that will either be useful to them or
help their colleague (i.e.~the presenter). Presenters should do their
best to satisfy those motivations, whether it is through the relevancy
of the concept or the data. It does not make sense to present a Code
Club on downloading stock market data if it is not useful or interesting
to the group. Similarly, participants should not be tasked with
improving the presenter's code if the presenter has no intention of
incorporating the suggestions. A list of packages or tasks that group
members are interested in along with a log of previous topics could help
a presenter struggling to find a topic choose one that will make a
rewarding Code Club.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. Beginners will benefit from sessions that cover fundamental
concepts and functions. The benefits experienced by more experienced
participants include the improved understanding of concepts that comes
through teaching the concepts and breaking down problems into simpler
elements. Furthermore, instead of focusing on core functions, the group
could balance Code Club sessions that cover basics with those that cover
methods and packages that are new to everyone in the group. We have also
identified several strategies to overcome the challenges presented by
participants at varying skill levels. Central to the Code Club format is
the use of paired-programming {[}15{]}. Instead of letting participants
form their own pairs, the presenter can select pairs of participants
with either similar or differing skill levels, depending on the
presenter's goals. Partnerships between those with similar skill levels
requires the presenter to design appropriate activities for each skill
level. We have found that commenting code is a good skill for beginners
since it forces them to dissect and understand a code chunk
line-by-line. It also reinforces the value of commenting as they develop
their skills and independence. An advantage of forming partnership
between people with disparate skill levels is that it is more likely for
groups to provide the presenter with a diverse range of methods that
achieve the same result. This approach to pairing also helps to graft
new members that have emerging programming skills. Regardless of how
partners are selected, consider asking the pairs to identify a navigator
and a driver {[}16{]}. The driver types at the computer while the
navigator tells them what to type, thus ensuring participation of both
partners. Midway through the activity, the presenter can have the
partners switch roles. Intentionally forming pairs can also engineer
group interactions by avoiding potentially disruptive partnerships or
pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion
about a paper that the participant has not read. In that context, coming
to Code Club without having installed a dependency is similar to asking
a simple question about the Journal Club paper without first reading it.
Both instances show a lack of preparation. Just as a presenter must
follow Rule 6 to provide materials ahead of time, participants must
review the code in advance, download the data sets, install the
necessary packages, and perhaps read up on the topic. If the Code Club
is based on a paper or chapters in a data science book (e.g.,
{[}17--19{]}), the participants should read them before the session and
consider how they might incorporate the concepts into their own work.

@@

\leavevmode\hypertarget{ref-Barone2017}{}%
2. Barone L, Williams J, Micklos D. Unmet needs for analyzing biological
big data: A survey of 704 NSF principal investigators. Ouellette F,
editor. PLOS Computational Biology. 2017;13: e1005755.
doi:\href{https://doi.org/10.1371/journal.pcbi.1005755}{10.1371/journal.pcbi.1005755}

\leavevmode\hypertarget{ref-Wilson2016}{}%
3. Wilson G. Software carpentry: Lessons learned. F1000Research. 2016.
doi:\href{https://doi.org/10.12688/f1000research.3-62.v2}{10.12688/f1000research.3-62.v2}

\leavevmode\hypertarget{ref-Feldon2017}{}%
4. Feldon DF, Jeong S, Peugh J, Roksa J, Maahs-Fladung C, Shenoy A, et
al. Null effects of boot camps and short-format training for PhD
students in life sciences. Proceedings of the National Academy of
Sciences. 2017;114: 9854--9858.
doi:\href{https://doi.org/10.1073/pnas.1705783114}{10.1073/pnas.1705783114}

\leavevmode\hypertarget{ref-Carpenter2012}{}%
5. Carpenter SK, Cepeda NJ, Rohrer D, Kang SHK, Pashler H. Using spacing
to enhance diverse forms of learning: Review of recent research and
implications for instruction. Educational Psychology Review. 2012;24:
369--378.
doi:\href{https://doi.org/10.1007/s10648-012-9205-z}{10.1007/s10648-012-9205-z}

\leavevmode\hypertarget{ref-Rohrer2015b}{}%
6. Rohrer D. Student instruction should be distributed over long time
periods. Educational Psychology Review. 2015;27: 635--643.
doi:\href{https://doi.org/10.1007/s10648-015-9332-4}{10.1007/s10648-015-9332-4}

\leavevmode\hypertarget{ref-Bud2010}{}%
7. Budé L, Imbos T, Wiel MW van de, Berger MP. The effect of distributed
practice on students' conceptual understanding of statistics. Higher
Education. 2010;62: 69--79.
doi:\href{https://doi.org/10.1007/s10734-010-9366-y}{10.1007/s10734-010-9366-y}

\leavevmode\hypertarget{ref-Karpicke2011}{}%
8. Karpicke JD, Blunt JR. Retrieval practice produces more learning than
elaborative studying with concept mapping. Science. 2011;331: 772--775.
doi:\href{https://doi.org/10.1126/science.1199327}{10.1126/science.1199327}

\leavevmode\hypertarget{ref-Rohrer2015a}{}%
9. Rohrer D, Dedrick RF, Stershic S. Interleaved practice improves
mathematics learning. Journal of Educational Psychology. 2015;107:
900--908.
doi:\href{https://doi.org/10.1037/edu0000001}{10.1037/edu0000001}

\leavevmode\hypertarget{ref-Karpicke2008}{}%
10. Karpicke JD, Roediger HL. The critical importance of retrieval for
learning. Science. 2008;319: 966--968.
doi:\href{https://doi.org/10.1126/science.1152408}{10.1126/science.1152408}

\leavevmode\hypertarget{ref-Lonsdale2016}{}%
11. Lonsdale A, Penington JS, Rice T, Walker M, Dashnow H. Ten simple
rules for a bioinformatics journal club. Lewitter F, editor. PLOS
Computational Biology. 2016;12: e1004526.
doi:\href{https://doi.org/10.1371/journal.pcbi.1004526}{10.1371/journal.pcbi.1004526}

\leavevmode\hypertarget{ref-Wilson2014}{}%
12. Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et al.
Best practices for scientific computing. Eisen JA, editor. PLoS Biology.
2014;12: e1001745.
doi:\href{https://doi.org/10.1371/journal.pbio.1001745}{10.1371/journal.pbio.1001745}

\leavevmode\hypertarget{ref-Kothiyal2014}{}%
13. Kothiyal A, Murthy S, Iyer S. Think-pair-share in a large CS1 class.
Proceedings of the 2014 conference on innovation \& technology in
computer science education - ITiCSE 14. ACM Press; 2014.
doi:\href{https://doi.org/10.1145/2591708.2591739}{10.1145/2591708.2591739}

\leavevmode\hypertarget{ref-StackOverflow_MREE}{}%
14. Stack Overflow. How to create a minimal, reproducible example. 2020.
Available:
\url{https://stackoverflow.com/help/minimal-reproducible-example}

\leavevmode\hypertarget{ref-Hanks2011}{}%
15. Hanks B, Fitzgerald S, McCauley R, Murphy L, Zander C. Pair
programming in education: A literature review. Computer Science
Education. 2011;21: 135--173.
doi:\href{https://doi.org/10.1080/08993408.2011.579808}{10.1080/08993408.2011.579808}

\leavevmode\hypertarget{ref-Bockeler_2020}{}%
16. Böckeler B, Siessegger N. On pair programming. 2020. Available:
\url{https://martinfowler.com/articles/on-pair-programming.html}

\leavevmode\hypertarget{ref-Wickham_2016}{}%
17. Wickham H, Grolemund G. O'Reilly Media; 2016. Available:
\url{https://r4ds.had.co.nz}

\leavevmode\hypertarget{ref-Schloss_2018}{}%
18. Schloss PD. The Riffomonas reproducible research tutorial series.
Journal of Open Source Education. 2018;1: 13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}

\leavevmode\hypertarget{ref-Noble2009}{}%
19. Noble WS. A quick guide to organizing computational biology
projects. Lewitter F, editor. PLoS Computational Biology. 2009;5:
e1000424.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000424}{10.1371/journal.pcbi.1000424}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.25\columnwidth}\raggedright
\textbf{Title}\strut
\end{minipage} & \begin{minipage}[b]{0.69\columnwidth}\raggedright
\textbf{Description}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.25\columnwidth}\raggedright
base vs.~ggplot2\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given input data and a figure, recreate the figure using R's base
graphics or ggplot2 syntax\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Snakemake\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a bash script that contains an analysis pipeline, convert it to a
Snakemake workflow
(\url{https://github.com/SchlossLab/snakemake_riffomonas_tutorial})\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
DRYing code\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given script with repeated code, create functions to remove
repetition\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Vegan R package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Compare microbial communities using the \texttt{adonis} function in the
Vegan R package
(\url{https://github.com/SchlossLab/Code_Review_42717})\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
tidy data\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a wide-formatted data table, convert it to a long, tidy-formatted
data table using tools from R's tidyverse\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
GitFlow\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Participants file and claim an issue to add their name to a README file
in a GitHub-hosted repository and file a pull request to complete the
issue\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
googledocs4 R package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Scrape a Google docs workbook and clean the data to identify previous
Code Club presenters\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Develop an R Package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Convert a lab member's collection of scripts into an R package over a
series of sessions\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Documenting R code\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Use roxygen2 to supplement comments in R code to improve documentation
(\url{https://github.com/SchlossLab/documenting-R})\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
gganimate R package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Convert static plots generated with ggplot2 package into GIFs
(\url{https://github.com/SchlossLab/2020-04-12-CodeClub_PlotAnimation/})\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}
